Distinguishing Spin-Aligned and Isotropic Black Hole Populations With Gravitational Waves

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #16, moving equations to methods section.

The spin parameter with the largest effect on waveforms, and a
correspondingly tight constraint from the data, is a mass-weighted,
aligned, ``effective spin,''
\begin{equation}
  \chieff = \frac{c}{GM} \left( \frac{\vec{S}_1}{m_1} + \frac{\vec{S}_2}{m_2}
  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} \equiv \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
\end{equation}
where $m_{1,2}$ are the gravitational masses of the more-massive (1)
and less-massive (2) components, $M = m_1 + m_2$ is the total mass,
$\vec{S}_{1,2}$ are the spin angular momentum vectors of the black
holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
assumed to point in the $\hat{z}$ direction, and
$0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
projections of the individual \ac{BH} spins\cite{2016PhRvL.116x1102A}.

@@

Each of our models for the merging \ac{BBH} spin population assumes
that the merging black holes are of equal mass (this is marginally
consistent with the three observations\cite{O1-BBH}, and the
$\chieff$ distribution is not particularly sensitive to the mass ratio
between the merging objects---see Section \ref{sec:mass-ratio}).  We
assume that the population spin distribution factorises into a
distribution for the spin magnitude $a$ and a distribution for the
spin angles.  Finally, we assume that the distribution of spins is
common to each component in a merging binary.  Choosing one of three
magnitude distributions
\begin{equation}
  \label{eq:magnitude-dists}
  p(a) = \begin{cases}
    2\left(1-a \right) & \textnormal{``low''} \\
    1 & \textnormal{``flat''} \\
    2 a & \textnormal{``high''}
  \end{cases},
\end{equation}
and pairing with either an isotropic angular distribution or a
distribution that generates perfect alignment with the positive $z$
axis yields six different models for the $\chieff$ distribution.
These models are shown in Figure \ref{fig:chieff-distribution-models}.

\begin{figure}
  \plotone{../plots/chi-eff-distributions}
  \caption{\label{fig:chieff-distribution-models} The models for the
    distribution of $\chieff$ considered in this paper.  In all models
    we assume that the binary mass ratio $q \equiv m_1/m_2 = 1$ and
    that the distribution of spin vectors is the same for each
    component.  The ``flat'' (blue lines), ``high,'' (green lines),
    and ``low'' (red lines) magnitude distributions are defined in
    Eq.\ \eqref{eq:magnitude-dists}.  Solid lines give the $\chieff$
    distribution under the assumption that the orientations of the
    spins are isotropic; dashed lines give the distribution under the
    assumption that both objects' spins are aligned with the orbital
    angular momentum.  The isotropic distributions are readily
    distinguished from the aligned distributions by the production of negative
    $\chieff$ values, while the distinction between the three models
    for the spin magnitude distribution is less sharp.}
\end{figure}

The distributions in Eq.\ \eqref{eq:magnitude-dists} are not meant to
represent any particular physical model, but rather to capture our
uncertainty about the spin magnitude distribution, discussed in detail
below; neither observations nor population synthesis codes can at this
point authoritatively suggest \emph{any} particular spin
distribution\cite{2015PhR...548....1M}.  Our models, however, allow us
to see how sensitive the $\chieff$ distribution is to spin alignment
given uncertainties about the spin magnitudes.

@@

\subsection{Future Spin Measurements}
\label{subsec:future}

@@

>>>>

The spin parameter with the largest effect on waveforms, and a
correspondingly tight constraint from the
data\cite{2016PhRvL.116x1102A}, is a mass-weighted combination of the
components of the dimensionless spin vectors of the two black holes
that are aligned with the orbital axis, the ``effective spin,''
$-1 \leq \chieff \leq 1$.

@@

Each of our models for the merging \ac{BBH} spin population assumes
that the merging black holes are of equal mass (this is marginally
consistent with the three observations\cite{O1-BBH}, and the $\chieff$
distribution is not particularly sensitive to the mass ratio between
the merging objects---see Section \ref{sec:mass-ratio}).  We assume
that the population spin distribution factorises into a distribution
for the spin magnitude $a$ and a distribution for the spin angles.
Finally, we assume that the distribution of spins is common to each
component in a merging binary.  Choosing one of three magnitude
distributions (see Methods Section
\ref{methsec:chieff-spin-magnitude}), ``low'' (mean $a = 0.33$,
standard deviation $0.24$), ``flat'' (mean $a = 0.5$, standard
deviation $0.29$), ``high'' (mean $a = 0.67$, standard deviation
$0.24$) and pairing with either an isotropic angular distribution or a
distribution that generates perfect alignment with the positive $z$
axis yields six different models for the $\chieff$ distribution.
These models are shown in Figure \ref{fig:chieff-distribution-models}.

\begin{figure}
  \plotone{../plots/chi-eff-distributions}
  \caption{\label{fig:chieff-distribution-models} The models for the
    population distribution of $\chieff$ considered in this paper.  In
    all models we assume that the binary mass ratio
    $q \equiv m_1/m_2 = 1$ and that the distribution of spin vectors
    is the same for each component.  The ``flat'' (blue lines),
    ``high,'' (green lines), and ``low'' (red lines) magnitude
    distributions are defined in Eq.\ \eqref{eq:magnitude-dists}.
    Solid lines give the $\chieff$ distribution under the assumption
    that the orientations of the spins are isotropic; dashed lines
    give the distribution under the assumption that both objects'
    spins are aligned with the orbital angular momentum.  The
    isotropic distributions are readily distinguished from the aligned
    distributions by the production of negative $\chieff$ values,
    while the distinction between the three models for the spin
    magnitude distribution is less sharp.}
\end{figure}

These magnitude distributions not meant to represent any particular
physical model, but rather to capture our uncertainty about the spin
magnitude distribution, discussed in detail below; neither
observations nor population synthesis codes can at this point
authoritatively suggest \emph{any} particular spin
distribution\cite{2015PhR...548....1M}.  Our models, however, allow us
to see how sensitive the $\chieff$ distribution is to spin alignment
given uncertainties about the spin magnitudes.

@@

@@

\section{Effective Spin and Spin Magnitude Distributions}
\label{methsec:chieff-spin-magnitude}

The effective spin is defined by\cite{2016PhRvL.116x1102A} 
\begin{equation}
  \chieff = \frac{c}{GM} \left( \frac{\vec{S}_1}{m_1} + \frac{\vec{S}_2}{m_2}
  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} \equiv \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
\end{equation}
where $m_{1,2}$ are the gravitational masses of the more-massive (1)
and less-massive (2) components, $M = m_1 + m_2$ is the total mass,
$\vec{S}_{1,2}$ are the spin angular momentum vectors of the black
holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
assumed to point in the $\hat{z}$ direction, and
$0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
projections of the individual \ac{BH} spins.

We form the population distributions of $\chieff$ shown in Figure
\ref{fig:chieff-distribution-models} by assuming that each black hole
in a binary has a dimensionless spin magnitude drawn from one of three
distributions,
\begin{equation}
  \label{eq:magnitude-dists}
  p(a) = \begin{cases}
    2\left(1-a \right) & \textnormal{``low''} \\
    1 & \textnormal{``flat''} \\
    2 a & \textnormal{``high''}
  \end{cases},
\end{equation}
referred to as ``low,'' ``flat,'' and ``high'' in the text above.
